Progress Report for CSS 586 Course Project: \\ Modeling Latent Patterns in Music

### 2021-05-01 papers/progress/alex/progress-alex.tex
pianoroll bar cropping

MuseGAN  \cite{dong2017musegan}.

@@

Several choices must be made in how to preprocess MIDI files into training
examples for a neural network. In order to accommodate polyphonic music, we will
convert each MIDI file into a piano roll representation, wherein each instrument
track is a sparse matrix of the velocity values for each of 128 possible pitch
levels at each timestep. Because songs are typically at least a few minutes long
and of widely varying length, we will crop songs into equal-length phrases.

The result of this preprocessing is that each training example will be a 3D
tensor of shape (tracks x notes x ticks).

>>>>

MuseGAN \cite{dong2017musegan} is an application of Generative Adversarial
Networks to polyphonic MIDI music generation, trained on four-bar phrases of a
multi-track pianoroll representation of rock songs from the Lakh Midi Dataset
\cite{raffel_learning-based_2016}.

@@

Several choices must be made in how to preprocess MIDI files into training
examples for a neural network. In order to accommodate polyphonic music, we will
convert each MIDI file into a pianoroll representation, wherein each instrument
track is a sparse matrix of the velocity values for each of 128 possible pitch
levels at each timestep. Because songs are typically at least a few minutes long
and of widely varying length, we will crop songs into equal-length phrases.

The result of this preprocessing is that each training example will be a 3D
tensor of shape (tracks x notes x ticks) and stacking the training examples will
produce a 4D tensor, which is also typically the dimension of image data
tensors in machine learning.

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{first_four_bars.png}
  \caption{The first four bars of Beethoven's Serioso String Quartet}
  \Description{A pianoroll representation of cello, viola and violin parts.}
\end{figure}

Data augmentation is also possible--the literature suggests augmentation via
pitch shifting the entire training example up or down by up to six semitones,
and increasing or reducing the speed by up to 10\% in order to create additional
training examples and reduce overfitting \cite{oore_this_2018}.
